### 2016-11-20 paper.tex
made some changes

\documentclass{IEEEtran}

\title{Access to food in Chicago: \\ a hierarchical perspective}
\author{Daniel Berry}

\begin{document}

\maketitle

\begin{abstract}
  Access to readily available and inexpensive healthy food options is important for personal health. 
\end{abstract}

\section{Introduction}

\section{Methods}

\section{Results}

\section{Conclusions}

\end{document}